### 2016-04-24 paper.tex
Replicated-ish proposal in TeX, file for paper, updated Makefile

\documentclass[letterpaper, 12pt]{article}
\input{default}

% Setup Chicago 16th edition, notes-mode citations
\usepackage[notes,short,backend=biber]{biblatex-chicago}
\bibliography{exported_items}

% Use English babel and fancy quotes
\usepackage[american]{babel}
\usepackage{csquotes}

\title{Medical Training for Abortion Procedures in America}
\author{Thomas R. Murphy}
\date{\today}

\begin{document}
\maketitle

	%Document Content
	Words to cite \autocite[pg. N]{turk_availability_2014}

\newpage

% Make biblography
\printbibliography

\end{document}

Post-document snippets

\cite[<prenote>][<postnote>]{<key>}
\cite[<postnote>]{<key>}