### 2016-04-29 references.tex
Create a file of all consulted references for this paper

\documentclass[letterpaper, 12pt]{article}
\input{default}

\usepackage[pdftex,
            pdfauthor={Thomas Russell Murphy},
            pdftitle={Medical Training for Abortion Procedures and Physician Provision of Abortions in the United States: References},
            pdfsubject={},
            pdfkeywords={abortion, training, medical training, OB/GYN, obstetrician, gynecology, reproduction, fertility, birth control},
            pdfproducer={\pdftexbanner},
            pdfcreator={{\pdftexbanner} with Biber and a Makefile}]{hyperref}

% Setup Chicago 16th edition, footnotes-mode citations
\usepackage[notes,short,backend=biber]{biblatex-chicago}
\bibliography{exported_items_active_updating}

% Use English babel and fancy quotes
\usepackage[american]{babel}
\usepackage{csquotes}

\title{Medical Training for Abortion Procedures and Physician Provision of Abortions in the United States: References}
\author{Thomas R. Murphy}
\date{\today\\ANTH 360}

\begin{document}
\maketitle

% Bibliographic material
\nocite{*}
\printbibliography

\end{document}

Post-document snippets

\cite[<prenote>][<postnote>]{<key>}
\cite[<postnote>]{<key>}